---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

Load the data from the csv file 
```{r echo = TRUE}
activity <- read.csv("activity.csv")
```

Convert the date variable to the date type
```{r echo = TRUE}
activity[,2] <- as.Date(activity[,2], "%Y-%m-%d")
```

Convert the interval variable to the hour.min 
```{r echo=TRUE}
activity[,3] <- activity[,3] %/% 100 + (activity[,3] %% 100) / 100
```


## What is mean total number of steps taken per day?

Let's ignore NA values for this part
```{r echo = TRUE} 
activity_no_na = na.omit(activity)
```

Load dplyr package and summarise activities for each day
```{r echo = TRUE} 
library(dplyr)
activity_per_day <- summarise(group_by(activity_no_na, date), steps = sum(steps))
```

Calculate a **histogram** of total number of steps taken per day.
```{r echo = TRUE} 
hist(activity_per_day$steps, 30, col = "red", main = "Histogram of total number of steps per day ",
     xlab = "Number of steps per day")
```

Calculate the **mean** of steps taken per day
```{r echo = TRUE} 
mean_without_na <- mean(activity_per_day$steps)
mean_without_na
```

Calculate the **median** of steps taken per day
```{r echo = TRUE} 
median_without_na <- median(activity_per_day$steps)
median_without_na
```


## What is the average daily activity pattern?
Summorise the interval activities and calculate the mean for each interval.
Plot the average daily activity pattern.
```{r echo=TRUE}
daily_pattern <- summarise(group_by(activity, interval), steps = mean(steps, na.rm = TRUE))
plot(daily_pattern$interval, daily_pattern$steps, type = 'l', col = "red", xlab = "Time of the day", 
     ylab = "Number of steps", main = "Average daily activity pattern")
```

Find the 5-min interval wich, on average, contains the **maximum number of steps**.
```{r echo=TRUE}
daily_pattern[which.max(daily_pattern$steps), "interval"]
```

## Imputing missing values
Calculate **the total number of missing values** 
```{r echo = TRUE}
nrow(activity[is.na(activity$steps),])
```

Correct the missing values with average value for this interval
```{r echo = TRUE}
activity_corrected <- activity
for (i in 1:nrow(activity_corrected)) {
  if(is.na(activity_corrected[i, 'steps'])){
    activity_corrected[i, 'steps'] <- daily_pattern[daily_pattern$interval == activity_corrected[i, 'interval'], 'steps']
  }
}
```

Summarise activities for each day for corrected dataset
```{r echo = TRUE} 
library(dplyr)
activity_per_day_corrected <- summarise(group_by(activity_corrected, date), steps = sum(steps))
```

Calculate a **histogram** of total number of steps taken per day for corrected dataset.
```{r echo = TRUE} 
hist(activity_per_day_corrected$steps, 30, col = "red", main = "Histogram of total number of steps per day (CORRECTED)",
     xlab = "Number of steps per day")
```

Let's look again on the histogram without NA
```{r echo = TRUE} 
hist(activity_per_day$steps, 30, col = "red", main = "Histogram of total number of steps per day (ORIGINAL)",
     xlab = "Number of steps per day")
```

Calculate the **mean** of steps taken per day for corrected dataset
```{r echo = TRUE} 
mean(activity_per_day_corrected$steps)
```
The mean without na:
```{r echo = TRUE} 
mean_without_na
```

Calculate the **median** of steps taken per day for corrected dataset
```{r echo = TRUE} 
median(activity_per_day_corrected$steps)
```
The median without na:
```{r echo = TRUE} 
median_without_na
```

As we can see, the values for mean and median total number of steps taken per day have not changed a lot.
The average stayed the same, the median changed a bit.


## Are there differences in activity patterns between weekdays and weekends?
